---
title: "Week 3 Exercises"
author: "J.Blondeau"
date: "2024-09-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All the following question use the sttst_all_sites.xlsx file in the /data directory.

All answers should be made in the .Rmd file, showing the code and your answer and everything pushed back up to your GitHub repo.

triple stars renders a thin line across the page

***
1. Import sttstj_all_sites and inspect the data set.  What is the data structure and the data type of each field?
The data has 4 columns 2 of which are numeric and 2 of wich are characters

```{r}
 #this is a code chunk, anything inside the back tics is like being in the R console.

library(readxl)
sttstj_all_sites <- read_excel("data/sttstj_all_sites.xlsx", 
    sheet = "locations")
head(sttstj_all_sites)

#glimpse(data)
```
***

2. How many samples are in the data set? 
There are 2505 samples in the data set.
```{r}
nrow(sttstj_all_sites)

 #print(nrow(data))
```

3. How many unique habitat types are there?  What are the names of each habitat?
There are 6 unique habitats. The names of each habitat are "PVMT" "AGRF" "SCR"  "PTRF" "BDRK" "HARD". 
```{r}
library(dplyr)
library(tidyr)

n_distinct(sttstj_all_sites$hab)
unique(sttstj_all_sites$hab)

# hab_types <- unique(data$hab) 
#print(hab_types)
```


4. How many sites are there in the "PVMT" habitat type?
There are 571 sites

```{r}
sttstj_all_sites %>%
  filter(hab == "PVMT")%>%
  summarize(length(unique(yr_site)))

#data %>% count(hab)

```

5. How many sites are there in the "PVMT" or "AGRF" habitat types?
There 1538 habitat types

```{r}
sttstj_all_sites %>%
  filter(hab == "PVMT"|hab == "AGRF")%>%
  summarize(length(unique(yr_site)))

#data %>% 
  # filter(hab %in% c("PVMT","AGRF")) %>% 
  # count(hab) %>% 
  # summarise(tot = sum(n))
```
6. How many sites are there in all the habitat types EXCEPT "PVMT"?

There are 1934 sites in all habitat types EXCEPT "PVMT"

```{r}
sttstj_all_sites %>%
  filter(hab != "PVMT") %>%
  summarize(length(unique(yr_site)))

# data %>% 
#   filter(hab != "PVMT") %>% 
#   count(hab) %>% 
#   summarise(tot = sum(n))

```

7. Show the first 5 records (rows) of...
    - all sites sampled in "AGRF" or "SCR"
    - only show the "yr_site" and "hab" columns
    
<!-- extra line break -->
<br>

```{r}
first_five <- sttstj_all_sites %>%
  filter(hab == "AGRF"|hab == "SCR") %>%
  select(yr_site,hab) %>%
  slice(1:5)

# data %>% 
#   filter(hab %in% c("AGRF", "SCR")) %>% 
#   head()
  
```

8. What is the northern most site sampled in the "BDRK" habitat?
The northern most site is 2019_5450.
```{r}
northern_most <- sttstj_all_sites %>%
  filter(hab == "BDRK") %>%
  arrange(desc(lat))
  
head(northern_most)

# data %>% 
#   filter(hab == "BDRK") %>% 
#   arrange(desc(lat))
```

9. What is the western most site sampled in either "AGRF" or "PTRF" habitat?
2015_9758 is the western most site sampled in either "AGRF" or "PTRF" habitat.
```{r}
western_most <- sttstj_all_sites %>%
  filter(hab == "AGRF" | hab == "PTRF") %>%
  arrange(lon)
  
head(western_most)

```
10. How many years were surveyed in the data set?
  16 years
```{r}

count_years <- sttstj_all_sites %>%
  separate(yr_site, c("year", "site", "_"))

count_years %>% 
  summarize(length(unique(year)))

# data %>% 
#   mutate(year = as.integer(substr(yr_site, 1,4))) %>% 
#   summarise(length(unique(year)))
```
11. What is the southern most site sample in the "PVMT" habitat in 2004?
2004_1932 is the southern most site sample in the "PVMT" habitat in 2004
```{r}
southern_most <- count_years %>%
  filter(hab == "PVMT", year == 2004) %>%
  arrange(lat)
  
head(southern_most)
```


